---
title: "Climate change and anomalies"
author: "Max"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

Let's look at some temperature anomalies, based on some data compiled by the NASA for the period 1951-1980.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}


library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)

library(infer)
library(wbstats)

```



# Climate change and temperature anomalies 



```{r weather_data, cache=TRUE}

weather <- read_csv("https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.csv", 
           skip = 1, 
           na = "***")

```




```{r tidyweather}

tidyweather<-weather %>%
  select(Year:Dec)%>%
 pivot_longer(!Year, names_to = "Month", values_to = "delta")

  
  
```



## Plotting Information



```{r scatter_plot}

tidyweather <- tidyweather %>%
  mutate(date = ymd(paste(as.character(Year), Month, "1")),
         month = month(date, label=TRUE),
         year = year(date))

ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies"
  )

```


```{r facet_wrap}
month.order <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
tidyweather$Month = factor(tidyweather$Month, levels = month.order)

ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies by month"
  )+
  facet_wrap(~Month)

```




```{r intervals}

comparison <- tidyweather %>% 
  filter(Year>= 1881) %>%     #remove years prior to 1881
  #create new variable 'interval', and assign values based on criteria below:
  mutate(interval = case_when(
    Year %in% c(1881:1920) ~ "1881-1920",
    Year %in% c(1921:1950) ~ "1921-1950",
    Year %in% c(1951:1980) ~ "1951-1980",
    Year %in% c(1981:2010) ~ "1981-2010",
    TRUE ~ "2011-present"
  ))

```



```{r density_plot}

ggplot(comparison, aes(x=delta, fill=interval))+
  geom_density(alpha=0.2) +   #density plot with tranparency set to 20%
  theme_bw() +                #theme
  labs (
    title = "Density Plot for Monthly Temperature Anomalies",
    y     = "Density"         #changing y-axis label to sentence case
  )

```

 

```{r averaging}

#creating yearly averages
average_annual_anomaly <- tidyweather %>% 
  group_by(Year) %>%   #grouping data by Year
  
  # creating summaries for mean delta 
  # use `na.rm=TRUE` to eliminate NA (not available) values 
  summarise(annual_average_delta = mean(delta, na.rm=TRUE)) 

#plotting the data:
ggplot(average_annual_anomaly, aes(x=Year, y= annual_average_delta))+
  geom_point()+
  
  #Fit the best fit line, using LOESS method
  geom_smooth() +
  
  #change to theme_bw() to have white background + black frame around plot
  theme_bw() +
  labs (
    title = "Average Yearly Anomaly",
    y     = "Average Annual Delta"
  )                         


```


## Confidence Interval for `delta`


```{r, calculate_CI_using_formula}

formula_ci <- comparison %>% 
  filter(interval == "2011-present")%>%
  

  # choose the interval 2011-present
  # what dplyr verb will you use? 

  # calculate summary statistics for temperature deviation (delta) 
  # calculate mean, SD, count, SE, lower/upper 95% CI
  # what dplyr verb will you use? 


   summarise(mean_delta = mean(delta , na.rm = TRUE), SD = sd(delta, na.rm = TRUE),count=n(), SE=SD/sqrt(count), lower_ci = mean_delta - 1.96 * SE, upper_ci = mean_delta + 1.96*SE)
 
 

#print out formula_CI
formula_ci
```


```{r, calculate_CI_using_bootstrap}

bootstrap <- comparison %>%
  filter(interval == "2011-present")%>%
  specify(response = delta)%>%
  generate(reps= 10000, type= "bootstrap")%>%
  calculate(stat = "mean")%>%
  get_ci(level=.95, type = "percentile")

bootstrap
```


From 2011 to present data the average change in the temperature is 0.966. After the filter function, we are left with 108 counts of data. There is a 95% chance of delta mean being in between 0.917 (lower confidence interval) and 1.02 (upper confidence interval). The deviation from the mean is 0.262. Standard error of 0.0252 indicates that the true population mean can be off by 0.0252 than the mean that we found.


