---
title: "Email Usage"
author: "Max"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

Let's look at how much time American people spend looking at their emails.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}


library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)

library(infer)
library(wbstats)

```


# General Social Survey (GSS)



```{r, read_gss_data, cache=TRUE}
gss <- read_csv(here::here("data", "smallgss2016.csv"), 
                na = c("", "Don't know",
                       "No answer", "Not applicable"))
```


## Email usage



```{r}
gss$emailmin<-as.numeric(gss$emailmin)
gss$emailhr<-as.numeric(gss$emailhr)
gss$email <- rowSums(cbind(gss$emailmin,gss$emailhr*60),na.rm=T)
gss =  gss %>%
  na.omit(emailmin)
```


```{r}

mean.email = mean(gss$email)
median.email= median(gss$email)
ggplot(gss, aes(x = email)) +
  geom_density(colour = "blue", fill = "lightblue")+
  geom_vline(xintercept = mean.email, size = 0.5, colour = "red") +       #add mean line
    geom_vline(xintercept = median.email, size = 0.5, colour = "green")+geom_text(aes(x=mean.email+270, label=paste0("Mean\n", round(mean.email,1)), y=0.0025)) + #add label to mean
    geom_text(aes(x=median.email, label=paste0("Median\n", median.email), y=0.0005)) + #add label to median
    theme_bw() +
    labs(title = "Most emails answered in 8 hours",
         x = "Minutes",
         y = "Density")
```


Median is a better measure of typical among of time Americans spend on email weekly since the density of median is higher than that of mean.


```{r}
bootstrap <- gss %>%
  specify(response = email)%>%
  generate(reps= 1000, type= "bootstrap")%>%
  calculate(stat = "mean")%>%
  get_ci(level=.95, type = "percentile")

bootstrap$lower_ci <- bootstrap$lower_ci/60
bootstrap$upper_ci <- bootstrap$upper_ci/60


lower_ci <- bootstrap %>%
  summarise (lower_ci, hours= floor(lower_ci))%>%
  mutate(minutes = (lower_ci-floor(lower_ci))*60)

upper_ci <- bootstrap %>%
  summarise (upper_ci, hours= floor(upper_ci))%>%
  mutate(minutes = (upper_ci-floor(upper_ci))*60)

lower_ci
upper_ci
```

We expect the confidence interval to be wider as we try to be more confident with the 99%. The wider the difference between the lower and upper confidence interval, the more sure we are.


