---
title: "Session 6: Homework 3"
author: "Pietro Gobbi"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---
Could physical activity and weight be two correlated variables among high schoolers ? Let's find out.

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```


# Youth Risk Behavior Surveillance

## Load the data

?yrbss

```{r}
data(yrbss)
glimpse(yrbss)
skim(yrbss)
```


## Exploratory Data Analysis

```{r, eda_on_weight}

Number_of_missing_weight <- count(yrbss,is.na(yrbss$weight))
Number_of_missing_weight[2,2]       ## take the number of missing weight only

temp <- yrbss%>%
  filter(weight != is.na(weight))     ## eliminate NAs to calculate summary statistics

mean.weight <- mean(temp$weight)
median.weight <- median(temp$weight)

summary <- yrbss %>%
  filter(!is.na(weight))%>%
  summarise(mean_weight = mean.weight , min_weight = min(weight), max = max(weight), median_weight = median.weight, sd_weights = sd(weights))
summary

ggplot(temp, aes(x= weight))+
  geom_density()+
  theme_bw()+
   geom_vline(xintercept = mean.weight, size = 0.5, colour = "red") +       #add mean line
    geom_vline(xintercept = median.weight, size = 0.5, colour = "green")+   #add median line
  geom_text(aes(x=mean.weight+8, label=paste0("Mean\n", round(mean.weight,2)), y=0.01))+
  geom_text(aes(x=median.weight-8, label=paste0("Median\n", round(median.weight,2)), y=0.02))+
  labs(title = "Distribution of weights", x="Weight", y="")
```

We are missing 1004 values for weight. Based on our graph, median weight of the participants is at 64.41kg, while mean weight is at 67.91kg.


```{r}
temp <- yrbss %>%
  filter(!is.na(physically_active_7d),!is.na(weight))%>%
  group_by(physically_active_7d)

temp$physically_active_7d <- as.character(temp$physically_active_7d)
  

ggplot(temp, aes(x=weight, y=physically_active_7d))+
  geom_boxplot()+
  theme_bw()+
  labs(title="Weight vs Physical activity", x="Weight", y="Number of days with physical act. in the last 7")


```
It seems that there is a positive correlation between weight and physical activity. High schooler's that had below 2 days of physical activity seem to have the lowest weight on average, while those that had 5 and above usually had higher weights. 

  
```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss %>% filter(!is.na(physical_3plus)) %>% 
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))


```

```{r}
physically_active_no = yrbss %>%
  filter(physical_3plus == "no")%>%
  summarise(count = n())

physically_active_total = yrbss %>%
  filter(!is.na(physical_3plus))%>%
  summarise(count = n())

prop.test(physically_active_no$count, physically_active_total$count)


```


```{r, boxplot}
temp <- yrbss%>%
  filter(!is.na(physical_3plus))

ggplot(temp, aes(x=weight, y=physical_3plus))+
  geom_boxplot()+
  theme_bw()+
  labs(title="Weight vs Physical activity", x="Weight", y="3 or more days with physical act. in the last 7?")
```
This boxplot confirms our previous considerations about the correlation between weight and physical activity. We notice that 25th percentile, median and 75th percentile are higher for the group that answered "Yes" to the question asked.
It is clear that high schoolers that had 3 or more days of physical activity in the last week have higher weight on average than those who did not. 

## Confidence Interval


```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight
            )

```

There is an observed difference of about 1.77kg (68.44 - 66.67), and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant.




